---
title: "Final Project"
author: "Matthew Bentham"
date: "2023-04-30"
output: html_document
css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center>
# Energy Demand Victoria 
##### MATH1318: Final Project \n\n
Matthew Bentham 3923076 \n
John Murrowood 3923075 \n
</center>

---

# 1. Introduction




*Note* : All code provided is written in Rstudio using R 4.2.0



```{r, include=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(TSA)
library(lmtest)
library(tseries)
library(forecast)
```

```{r, include=FALSE}
sort.score <- function(x, score = c("bic", "aic")){
if (score == "aic"){
x[with(x, order(AIC)),]
} else if (score == "bic") {
x[with(x, order(BIC)),]
} else {
warning('score = "x" only accepts valid arguments ("aic","bic")')
}
}
```


# 2. Preliminary Analysis

## 2.1 Import and Analyse Dataset

To further investigate the time series data, the data set is first imported and converted to a time series object. This time series object is then plotted using the plot() function in order to visualize its main characteristics.

```{r , message=FALSE,fig.align="center",class.source="cobalt"}

Data <- read_csv("complete_dataset.csv")
Data


```
\n 
